---
title: "Risk Prediction"
author: "Serdar Turkarslan"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(caret)
library(pROC)
library(ggplot2)
library(tidyverse)

```


# Risk predicion from Carl
## Apply elastic net regression
## libraries: caret, pROC, ggplot2
## data - data.frame of network activity, rows are features, columns are samples
## classer - vector of risk labels (high, low). must match column order of data
## ml_method - name of caret ml method
## tuner - number of combinations of tuning parameter
## seed - random seed
## prune - if true remove low info features
## prune_cut - threshold for pruning (minimum number of absolute active features necessary)
## boot_iters - number of bootstrapping iterations

### Load data disease relevant
```{r load.data}
program_data <- read_csv("data/program_activity_disease_relevant.csv") 
TCGA_Survival_Gbm <- read_csv("data/TCGA_Survival_Gbm.csv") 



rownames(program_data) <- program_data$X1
xs <- intersect(TCGA_Survival_Gbm$Patient_ID, colnames(program_data))
diff <- setdiff(colnames(program_data), TCGA_Survival_Gbm$Patient_ID)

program_data_filt <- program_data %>%
  select(matches(xs)) %>%
  select(!matches(diff))

classer <- TCGA_Survival_Gbm %>%
  filter(Patient_ID %in% xs) %>%
  pull(Classer)

```

### Elastic net function
```{r elcticnet.function}
applyElasticNet <- function(data, classer, ml_method="glmnet", tuner=20, seed=123,
                                 prune=TRUE, prune_cut=3, boot_iters=1000) {
 
  pro <- data.frame(t(as.matrix(data)))
  colnames(pro) <- as.character(rownames(data))
  input <- pro
 
  classer <- ifelse(classer=="low", "low", "high")
 
  ## prune low info programs
  if(prune) {
     act <- apply(pro, 2, function(x) return(sum(abs(x))))
     input <- pro[,act >prune_cut]
  }
 
  print(paste("Model consists of", ncol(input), "programs"))
  input <- cbind(risk=classer, input)
 
  set.seed(seed)
  model <- caret::train(risk~., data=input, method = ml_method,
    trControl = trainControl("boot", number = boot_iters),
    tuneLength = tuner)
 
  ## get predictions
  preds <- predict(model, input[,-1], type="prob")
  out_preds <- data.frame(preds=preds[,1], risk=classer, sample=rownames(input))
 
  ## get predicted groups
  pred_group <- rep("low", nrow(preds))
  pred_group[preds[,1] > 0.5] <- "high"
  pred_group[preds[,1] > 0.6] <- "extreme"
 
 
  rocc <- roc(response=classer, predictor=preds[,1], plot=TRUE)
  auc1 <-  signif(auc(rocc)[1], 3)
  roc_d1 <- data.frame(FPR=rev(1-rocc$specificities), TPR=rev(rocc$sensitivities))
 
  p <- ggplot(roc_d1,aes(FPR,TPR))+geom_line(size = 2, alpha = 0.7) +
    labs(title= "Validation data ROC",
         x = "False Positive Rate",
         y = "True Positive Rate") +
    annotate('text', x=0.05, y=1, label="AUC", color="black", fontface = 'bold') +
    annotate('text', x=0.05, y=0.95, label=auc1, color="black", fontface = 'bold')  +
    geom_abline(slope=1, intercept=0, linetype="dashed")
 print(p)
  return(list(model, rocc, out_preds, pred_group))
}
```


```{r disease_rel_program_risk_prediction}

disease_rel_risk <- applyElasticNet(data = program_data_filt, classer = classer)
```


### Load data all
```{r load.data.all}
program_data <- read_csv("data/program_activity_all.csv") 
TCGA_Survival_Gbm <- read_csv("data/TCGA_Survival_Gbm.csv") 


rownames(program_data) <- program_data$X1
xs <- intersect(TCGA_Survival_Gbm$Patient_ID, colnames(program_data))
diff <- setdiff(colnames(program_data), TCGA_Survival_Gbm$Patient_ID)

program_data_filt <- program_data %>%
  select(matches(xs)) %>%
  select(!matches(diff))

classer <- TCGA_Survival_Gbm %>%
  filter(Patient_ID %in% xs) %>%
  pull(Classer)

```



```{r all_program_risk_prediction}

all_prog_risk <- applyElasticNet(data = program_data_filt, classer = classer)
```
